Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)

### 2022-09-22 paper.tex
Update Reflow
Update Introduction
Work in progress.

\section{Introduction}
\label{sec:1}

Bio-inspired algorithms have been very successful when used to solve complex
optimization problems
\cite{castillo2019comparative,valdez2021swarm,acherjee2020ultrasonic}, but as
their complexity increases so does the computing power required
\cite{ontiveros2018high}. One strategy to address this processing need is to
use distributed computing \cite{thain2005distributed}, or the resources
available in the cloud \cite{garcia2013there,eshratifar2019bottlenet} to help
find the solution. This strategy gives us elasticity to adjust (increase or
reduce) the computing power to achieve a balance according to the nature of the
problem.

One of the problems that we have identified in this research is that most
bio-inspired algorithms are designed with a traditionally sequential
perspective \cite{porto2018evolutionary,back1996evolutionary}, where each
process must wait for the previous task to finish before continuing. Some
architectures address this issue
\cite{valdez2021container,garcia2015evospace,merelo2016nodio}, what
distinguishes this proposal is the introduction of  an algorithm designed in its
entirety as a native solution in the cloud, fully distributed, where its
processes are executed in parallel and interaction takes place asynchronously. This technique makes it
easy to scale the computing power according to the complexity required by the
problem \cite{armbrust2010view}.

We designed our algorithm using observation, analysis, and abstraction in
nature to identify what most species in the animal kingdom have in common,
where individuals of a population that best adapt to the environment have
greater chances of survival, reproduction, and improvement of the species.
Imagining as if the task of writing these rules of evolution was in our hands,
questioning how we could solve it?

Like nature, we identified what most species have in common in the life cycle.
Our algorithm works on a constantly evolving population, experiencing the
stages that all living beings undergo \cite{read1968system}: being born,
growing up, reproducing, and dying, where each of these stages will be
processes (of our algorithm) that randomly affect the population individuals,
emulating the organic way.

The main contribution of the paper is to design and create a new algorithm with
the purpose to demonstrate that it is possible to evolve a population of
individuals, similar to a Genetic Algorithm (GA), using a distributed,
parallel, and asynchronous methodology by implementation and testing of the
algorithm.

We present the description of the following sections in our paper. First, we
introduce our algorithm inspired by the Life-Cycle of animal species in
% Please use capitalization consistently; I don't think this should go in Caps - JJ
section~\ref{section.proposal}, followed by our experimental configuration and
results in section~\ref{section.experiments}, we analyze and describe some of
our research findings in section~\ref{section.discussion}, and we finalize by
presenting some inferences based on our experiments' results in
section~\ref{section.conclusions}.

>>>>

\newpage
\section{Introduction}
\label{sec:1}

Bio-inspired algorithms have been very successful when used to solve complex
optimization problems
\cite{castillo2019comparative,valdez2021swarm,acherjee2020ultrasonic}, but as
their complexity increases so does the computing power required to execute them
\cite{ontiveros2018high}. One strategy to address this processing need is to
use distributed computing \cite{thain2005distributed}, or the resources
available in the cloud to help find the solution
\cite{garcia2013there,eshratifar2019bottlenet}. This strategy gives us
elasticity to increase (or reduce) the computing power to achieve a balance
according to the nature of the problem.

Most bio-inspired algorithms are traditionally designed with a sequential
(synchronous) perspective \cite{porto2018evolutionary,back1996evolutionary},
where each process must wait for the previous task to finish before continuing.
Our proposal presents an algorithm in its entirety as a native solution in the
cloud, fully distributed, where its processes are executed in parallel and
asynchronously. We designed the algorithm as a cloud-native solution using the
cloud available resources to divide the processing workload, among several
computers or running the algorithm as a cloud service. This technique makes it
easy to scale the computing power according to the complexity required by the
problem \cite{armbrust2010view}.

In previous work, we introduced an algorithm inspired by the biological
life-cycle of animal species. As in nature, population individuals grow older
and age. In this algorithm's reproduction process, couples must match by mutual
attraction, where sometimes individuals won't procreate offspring because of
the mate's low appeal. As time passes, the environment kills its individuals
either because of low fitness or age, causing occasional population extinction.
When this condition presents itself, it is required to create a new population
to continue evolution (restart). We previously used to generate a new random
set of solutions, with the cost of losing all evolution knowledge.

The main contribution of this paper is to demonstrate that it is possible to
evolve a population of individuals, similar to a Genetic Algorithm (GA), using
a distributed, parallel, and asynchronous methodology by the algorithm
implementation and testing. We improve the algorithm behaviour by solving
population extinction caused by nature's pressure in the Animal Life Cycle
Algorithm with one of the most promising alternatives: the projection of the
Elite towards the Champion, where we compare the results obtained with the
latter and the random use of the more traditional alternatives (mutation and
crossover), using classic benchmark functions for optimization for comparison.

We present the description of the following sections in our paper. First, we
illustrate our algorithm model, the problem found with the occasional
population extinction and our proposed solution in
section~\ref{section.proposal}, followed by our experimental configuration and
results in section~\ref{section.experiments}, where we continue to analyze and
describe some of our research findings in section~\ref{section.discussion}. We
finalize by presenting some inferences based on our experiments results in
section~\ref{section.conclusions}.

%PREVIOUSLY / FROM THE FIRST CHAPTER
%We designed our algorithm using observation, analysis, and abstraction in nature to identify what most species in the animal kingdom have in common, where individuals of a population that best adapt to the environment have greater chances of survival, reproduction, and improvement of the species. Imagining as if the task of writing these rules of evolution was in our hands, questioning how we could solve it?
%Like nature, we identified what most species have in common in the life cycle. Our algorithm works on a constantly evolving population, experiencing the stages that all living beings undergo \cite{read1968system}: being born, growing up, reproducing, and dying, where each of these stages will be processes (of our algorithm) that randomly affect the population individuals, emulating the organic way.
